---
title: "Assignment 4"
author: "Sneha Kumar and Temesgen"
date: "`r Sys.Date()`"
output: html_document
---
# Assignment 4

For this assignment you will write a new Quartro document based on the analysis of power required for different *possible_cars* from looping.RMD; You will

Load Function
```{r}
library(here)
source("~/ESM262/ESM262/autopower.R")
```
Add an additional super light car with mass 5,000 kg and a surface area 10m2
```{r}
possible_cars <- data.frame(name=c("A","B","C","New"), mass=c(10000,65000,38000,5000), area=c(22,30,22,10),power=autopower(V=28, A=possible_cars$area, m=possible_cars$mass))
print(possible_cars)
```
Compute the mean and maximum power for Cars A, B, C and for your new car, assuming that average highway speed is 80 km/hr with a standard deviation of 10km/hr
```{r}
nsample = 200
meanspeed_meters = 80*0.277
speeds = rnorm(mean=meanspeed_meters, sd=10, nsample)
summary(speeds)
```
The mean speed is 22.812 km/hr and the max speed is 53.933 km/hr.
```{r}
possible_cars$maxpower = autopower(V=53.933, A=possible_cars$area, m=possible_cars$mass)
possible_cars$meanpower = autopower(V=22.812, A=possible_cars$area, m=possible_cars$mass)
print(possible_cars)
```

Use 2 different methods to compute the mean and maximum power for each car type
a) **FOR** loop
```{r}
df_forloop <- as.data.frame(matrix(nrow=length(speeds), ncol=nrow(possible_cars)))
library(dplyr)
df_forloop <- df_forloop |>
  rename(A=V1, B=V2, C=V3, New=V4)

for (i in 1:ncol(df_forloop)) {
  df_forloop[,i] = autopower(A=possible_cars$area[i], m=possible_cars$mass[i], V=speeds)
}

view(df_forloop)
```

```{r}
meanvalues_forloop <- sapply(df_forloop, mean)
maxvalues_forloop <- sapply(df_forloop,max)
print(meanvalues_forloop)
print(maxvalues_forloop)
```
b) the **pmap** function from **purr**
```{r}
library(purrr)
df_pmap <- as.data.frame(pmap(list(A = possible_cars$area, m=possible_cars$mass), autopower, V=speeds))

colnames(df_pmap) <- c("A", "B", "C", "New")

view(df_pmap)

meanvalues_pmap <- sapply(df_pmap, mean)
maxvalues_pmap <- sapply(df_pmap,max)
print(meanvalues_pmap)
print(maxvalues_pmap)
```
Create two boxplots (one for each method (**FOR** and **pmap**)) that show the range of power consumption (across sampled speeds) for each car type.
```{r}
forlooppivot <- df_forloop |>
  pivot_longer(cols = everything(), names_to = "type", values_to = "power")
forlooppivot

library(ggplot2)
plot_forloop <- ggplot(data = forlooppivot) +
  geom_boxplot(aes(type, power, fill=type)) + 
  ggtitle("For Loop Boxplots")
view(plot_forloop)
```

Put the Quatro in your assignment github and submit a link on Canvas for Assignment 4 when its ready to be graded

