---
title: "Assignment 5"
author: "Sneha Kumar, Bella Perez, and Kyle Alves"
format: pdf
editor: visual
---

```{undefined}
#| label: Load Libraries
#| include: false
library(purrr)
library(ggplot2)
library(tidyverse)
library(testthat)
library(here)
```

Part 1: Design a Function

Our function is called "forest_sequestration," and iss sourced below.

Step 1: Create Data

We created a data frame of 500 observations, with a normal distribution of tree diameter and tree height

```{r}
treeID <- c(1:500)
tree_data <- data.frame(treeID,
                        diameter = rnorm(500, mean=22, sd=5),
                        height = rnorm(500, mean=15, sd=3),
                        tree_dimensions = NA,
                        CO2_sequestration = NA
                        )
View(tree_data)

source(here::here("Function", "forest_sequestration.R"))
```

Part 2: Loops

We used *for* loops to populate the two output variables of the function and add them to the original data frame.

```{r}
for(i in 1:nrow(tree_data)){tree_data$tree_dimensions = tree_data$diameter*tree_data$height}

for(i in tree_data$treeID[i]) {tree_data$CO2_sequestration = forest_sequestration(
                                                    diameter = tree_data$diameter,
                                                    height= tree_data$height)}
View(tree_data)
```

We then used the pmap function to do the same, and created a new data frame of forest sequestration.

```{r}
library(purrr)
#running function for each value in parameters with pmap
#turn into dataframe
ftsn2 <- as.data.frame(pmap(list(diameter = tree_data$diameter,
                                    height = tree_data$height),
                               forest_sequestration))
colnames(ftsn2) <- c(1:500)
View(ftsn2)


ftsn2_longer <- pivot_longer(ftsn2, cols = everything(), names_to = "treeID", values_to = "FtSn")

View(ftsn2_longer)
```

The two sets of data have matching values for each tree observation.

Part 3: Graph

We graphed the tree dimension values by the carbon sequestration values, to show how increasing tree volumes leads to increasing amounts of carbon sequestration in the forest. The scatter plot below shows this relationship. The box plots show the distribution of tree volumes and carbon sequestration values, to better understand the average values and range for both variables.

```{r}
plot1 <- ggplot(data=tree_data,aes(x=tree_dimensions, y=CO2_sequestration)) +
  geom_point() +
  xlab("Average Tree Dimension (Volume of Tree)") +
  ylab("Carbon Sequestration of the Forest")
plot1

library(ggExtra)

plot2 <- ggMarginal(plot1, type="boxplot")
plot2
```

Part 4: Testing

Our testing functions are stored in the file "ftsn_testing.R," sourced below:

```{r}
source(here::"testing/ftsn_testing.R")
```
