---
title: "Assignment 5"
author: "Sneha, Bella, and Kyle"
format: pdf
editor: visual
---

Design a function - you can pick any subject and you can even make up the equations as long as they conceptually make sense Make sure it has at least 2 inputs and 1 parameter (ideally more) and at least 2 outputs Code your function in R, save as a R file in subdirectory called R Make sure you include documentation (both at the top as we've shown in past examples, and inline). Include some error checking

The function created is called "function_name.r"

In an Quatro file, generate some data for 2 of the function inputs

```{r}

# If raw tree data is provided by user, calculate the averages and density.
  if (!is.null(tree_data)) {
    avg_diameter <- mean(tree_data$diameter, na.rm = TRUE)
    avg_height   <- mean(tree_data$height, na.rm = TRUE)
  
  # Combine tree dimensions into single variable for clarity / modularity
  tree_dimensions <- avg_diameter * avg_height
  
  # Estimate biomass per tree based on combined tree dimensions
  biomass_per_tree <- bm_conversion_factor * tree_dimensions
  
  # Total biomass per hectare (in tons)
  total_biomass <- tree_density * biomass_per_tree
  
  # Annual forest carbon sequestration: (growth-related biomass increase) * (carbon fraction)
  annual_sequestration <- total_biomass * growth_rate * carbon_fraction
  
  # Return both outputs as a list
  return(list(annual_sequestration = annual_sequestration, total_biomass = total_biomass))

```

Question 1: use a *for* loop to run the function for the data

```{r}

```

Question 2: repeat the "looping" using something from the "purrr" package

```{r}

```

Question 3: Graph results (you can decide what the most interesting way to graph - you just need to make one graph)

```{r}

```

Question 4: Write at least 2 tests for your function; store in a separate test file

```{r}

```

Put this in a git repo that includes an R subdirectory, a tests subsdirectory and your Quatro file and submit the link on Canvas (you can put all of this in a new git repo or a subdirectory of an existing git repo).

Submit the link to repo on Canvas (and name of subdirectory if part of a larger git project) as a group with your partner,
