---
title: "Assignment 4"
author: "Sneha Kumar and Temesgen"
date: "`r Sys.Date()`"
output: html_document
---
# Assignment 4

For this assignment you will write a new Quartro document based on the analysis of power required for different *possible_cars* from looping.RMD; You will


Load Function
```{r}
library(here)
source("~/ESM262/ESM262/autopower.R")
```
Add an additional super light car with mass 5,000 kg and a surface area 10m2
```{r}
possible_cars = data.frame(name=c("A","B","C","New"), mass=c(10000,65000,38000,5000), area=c(22,30,22,10),power=autopower(V=28, A=possible_cars$area, m=possible_cars$mass))
print(possible_cars)

```
Compute the mean and maximum power for Cars A, B, C and for your new car, assuming that average highway speed is 80 km/hr with a standard deviation of 10km/hr
Use 2 different methods to compute the mean and maximum power for each car type
a) **FOR** loop
```{r}

```
b) the **pmap** function from **purr**
```{r}

```
Create two boxplots (one for each method (**FOR** and **pmap**)) that show the range of power consumption (across sampled speeds) for each car type.

Put the Quatro in your assignment github and submit a link on Canvas for Assignment 4 when its ready to be graded

